Projecting genetic associations through gene expression patterns highlights disease etiology and drug mechanisms

### 2023-06-17 output/latex/manuscript.tex
add supplementary references section

We assessed our GLS model type I error rates (proportion of \(p\)-values below 0.05) and calibration using a null model of random traits and genotype data from 1000 Genomes Phase III.
We selected 312 individuals with European ancestry, and then analyzed 1,000 traits drawn from a standard normal distribution \(\mathcal{N}(0,1)\).
We ran all the standard procedures for the TWAS approaches (S-PrediXcan and S-MultiXcan), including:
1) a standard GWAS using linear regression under an additive genetic model,
2) different GWAS processing steps, including harmonization and imputation procedures as defined in\textsuperscript{\protect\hyperlink{ref-SiobXsoB}{116}},
3) S-PrediXcan and S-MultiXcan analyses.
Below we provide details for each of these steps.

@@

\textbf{Step 2 - GWAS processing}. These steps include harmonization of GWAS and imputation of \(z\)-scores, which are part of the TWAS pipeline and are needed in order to ensure an acceptable overlap with SNPs in prediction models.
The scripts to run these steps are available in\textsuperscript{\protect\hyperlink{ref-GPHGnFRN}{117}}.
These procedures were run for all 1,000 random phenotypes and generated a total number of 8,325,729 variants, including those with original and imputed \(z\)-scores.

@@

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched (using FGSEA\textsuperscript{\protect\hyperlink{ref-Z8WXLD67}{51}}) for the lipids-increasing gene-set from the CRISPR-screen (unadjusted \(p\)-values \textless{} 0.01).
LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_increase}}\label{tbl:sup:lipids_crispr:modules_enriched_increase}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
\textbf{LV246} & increase & \emph{DGAT2}, \emph{ACACA} & 0.0035 \\
LV702 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0046 \\
\textbf{LV607} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0058 \\
LV890 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0067 \\
\textbf{LV74} & increase & \emph{MBTPS1}, \emph{DGAT2} & 0.0078 \\
\textbf{LV865} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0092 \\
LV841 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0096 \\
\bottomrule()
\end{longtable}

@@

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched (using FGSEA\textsuperscript{\protect\hyperlink{ref-Z8WXLD67}{51}}) for the lipids-decreasing gene-set from the CRISPR-screen (unadjusted \(p\)-values \textless{} 0.01).
LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_decrease}}\label{tbl:sup:lipids_crispr:modules_enriched_decrease}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
LV520 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0006 \\
LV801 & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0022 \\
LV512 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0025 \\
\textbf{LV612} & decrease & \emph{PTEN}, \emph{FBXW7} & 0.0036 \\
LV41 & decrease & \emph{PCYT2}, \emph{TCF7L2} & 0.0041 \\
\textbf{LV838} & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0070 \\
LV302 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0083 \\
LV959 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0092 \\
\bottomrule()
\end{longtable}

@@

\begin{figure}
\hypertarget{fig:sup:lv136}{%
\centering
\includesvg[width=0.8\textwidth,height=\textheight]{images/lvs_analysis/lv136/lv136-cell_types.svg}
\caption{\textbf{Cell types for LV136.}
Pulmonary microvascular endothelial cells were exposed to hypoxia for 24 hours or more\textsuperscript{\protect\hyperlink{ref-VKYS05n1}{118}};}\label{fig:sup:lv136}
}
\end{figure}

@@

>>>>

We assessed our GLS model type I error rates (proportion of \(p\)-values below 0.05) and calibration using a null model of random traits and genotype data from 1000 Genomes Phase III.
We selected 312 individuals with European ancestry, and then analyzed 1,000 traits drawn from a standard normal distribution \(\mathcal{N}(0,1)\).
We ran all the standard procedures for the TWAS approaches (S-PrediXcan and S-MultiXcan), including:
1) a standard GWAS using linear regression under an additive genetic model,
2) different GWAS processing steps, including harmonization and imputation procedures as defined in\textsuperscript{\protect\hyperlink{ref-SiobXsoB}{1}},
3) S-PrediXcan and S-MultiXcan analyses.
Below we provide details for each of these steps.

@@

\textbf{Step 2 - GWAS processing}. These steps include harmonization of GWAS and imputation of \(z\)-scores, which are part of the TWAS pipeline and are needed in order to ensure an acceptable overlap with SNPs in prediction models.
The scripts to run these steps are available in\textsuperscript{\protect\hyperlink{ref-GPHGnFRN}{2}}.
These procedures were run for all 1,000 random phenotypes and generated a total number of 8,325,729 variants, including those with original and imputed \(z\)-scores.

@@

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched (using FGSEA\textsuperscript{\protect\hyperlink{ref-Z8WXLD67}{3}}) for the lipids-increasing gene-set from the CRISPR-screen (unadjusted \(p\)-values \textless{} 0.01).
LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_increase}}\label{tbl:sup:lipids_crispr:modules_enriched_increase}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
\textbf{LV246} & increase & \emph{DGAT2}, \emph{ACACA} & 0.0035 \\
LV702 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0046 \\
\textbf{LV607} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0058 \\
LV890 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0067 \\
\textbf{LV74} & increase & \emph{MBTPS1}, \emph{DGAT2} & 0.0078 \\
\textbf{LV865} & increase & \emph{ACACA}, \emph{DGAT2} & 0.0092 \\
LV841 & increase & \emph{ACACA}, \emph{DGAT2} & 0.0096 \\
\bottomrule()
\end{longtable}

@@

\begin{longtable}[]{@{}llll@{}}
\caption{Gene modules (LVs) nominally enriched (using FGSEA\textsuperscript{\protect\hyperlink{ref-Z8WXLD67}{3}}) for the lipids-decreasing gene-set from the CRISPR-screen (unadjusted \(p\)-values \textless{} 0.01).
LVs significantly aligned with pathways (FDR \textless{} 0.05) from the MultiPLIER models are shown in boldface.
\label{tbl:sup:lipids_crispr:modules_enriched_decrease}}\label{tbl:sup:lipids_crispr:modules_enriched_decrease}\tabularnewline
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endfirsthead
\toprule()
Gene module & Lipids gene-set & Leading edge & p-value \\
\midrule()
\endhead
LV520 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0006 \\
LV801 & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0022 \\
LV512 & decrease & \emph{FBXW7}, \emph{TCF7L2} & 0.0025 \\
\textbf{LV612} & decrease & \emph{PTEN}, \emph{FBXW7} & 0.0036 \\
LV41 & decrease & \emph{PCYT2}, \emph{TCF7L2} & 0.0041 \\
\textbf{LV838} & decrease & \emph{UBE2J2}, \emph{TCF7L2} & 0.0070 \\
LV302 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0083 \\
LV959 & decrease & \emph{TCF7L2}, \emph{PTEN} & 0.0092 \\
\bottomrule()
\end{longtable}

@@

\begin{figure}
\hypertarget{fig:sup:lv136}{%
\centering
\includesvg[width=0.8\textwidth,height=\textheight]{images/lvs_analysis/lv136/lv136-cell_types.svg}
\caption{\textbf{Cell types for LV136.}
Pulmonary microvascular endothelial cells were exposed to hypoxia for 24 hours or more\textsuperscript{\protect\hyperlink{ref-VKYS05n1}{4}};}\label{fig:sup:lv136}
}
\end{figure}

@@

\hypertarget{supplementary_references}{%
\subsection{Supplementary References}\label{supplementary_references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-SiobXsoB}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{Barbeira, A. N. \emph{et al.} \href{https://doi.org/10.1002/gepi.22346}{Fine‐mapping and QTL tissue‐sharing information improves the reliability of causal gene identification}. \emph{Genetic Epidemiology} \textbf{44}, 854--867 (2020).}

\leavevmode\vadjust pre{\hypertarget{ref-GPHGnFRN}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\emph{\href{https://github.com/hakyimlab/summary-gwas-imputation}{Harmonization and Imputation Overview}}. (hakyimlab, 2022).}

\leavevmode\vadjust pre{\hypertarget{ref-Z8WXLD67}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Korotkevich, G. \emph{et al.} Fast gene set enrichment analysis. (2016) doi:\href{https://doi.org/10.1101/060012}{10.1101/060012}.}

\leavevmode\vadjust pre{\hypertarget{ref-VKYS05n1}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Homo sapiens (ID 232177) - BioProject - NCBI. \url{https://www.ncbi.nlm.nih.gov/bioproject/PRJNA232177}.}
